### 2020-07-22 include/bab5.tex
22 7

\chapter{KESIMPULAN DAN SARAN}

\section{Kesimpulan}
% \subsection{Citra Digital}
\blindtext

\section{Saran}
\blindtext