### 2021-01-24 include/lampiran.tex
.

\chapter*{LAMPIRAN}

\blindtext

% \subsection*{Percobaan 1}

\begin{small}
    % \caption{Perbandingan FPS}
    \label{table:arm-aveblur1}
    \csvreader[
        % column count = 11,
        head to column names,
        longtable=cccccccc,
        separator=semicolon,
        before table=\rowcolors{2}{gray!15}{gray!30},
        table head= \rowcolor{gray!50!black} 
            \color{white} PID & 
            \color{white} VIRT & 
            \color{white} RES &
            \color{white} SHR & 
            \color{white} Status &  
            \color{white} CPU(\%) &
            \color{white} MEM(\%) & 
            \color{white} Time(s) 
            \\]
        {tables/performance-test/arm-aveblur5.csv}
        {
            PID=\PID, 
            VIRT=\VIRT, 
            RES=\RES,
            SHR=\SHR, 
            S=\S,
            CPU=\CPU,
            MEM=\MEM, 
            TIME=\TIME}
        {
            \PID & 
            \VIRT & 
            \RES &
            \SHR & 
            \S &
            \CPU &
            \MEM & 
            \TIME}
\end{small}