### 2021-01-29 include/halaman-abstrak.tex
30 01

\chapter*{ABSTRAK}

% latar belakang
% metode
Filter spasial linear
% hasil
% simpulan