### 2021-02-01 include/lampiran.tex
inc

\chapter*{LAMPIRAN}

\blindtext

% \subsection*{Percobaan 1}

\begin{table}
    \begin{small}
        \caption{ARM Average Blur Percobaan 1}
        \label{table:arm-aveblur1}
        \csvreader[
            % column count = 11,
            head to column names,
            longtable=cccccccc,
            separator=semicolon,
            before table=\rowcolors{2}{gray!15}{gray!30},
            table head= \rowcolor{gray!50!black} 
                \color{white} PID & 
                \color{white} VIRT & 
                \color{white} RES &
                \color{white} SHR & 
                \color{white} Status &  
                \color{white} CPU(\%) &
                \color{white} MEM(\%) & 
                \color{white} Time(s) 
                \\]
            {tables/performance-test/arm-aveblur5.csv}
            {
                PID=\PID, 
                VIRT=\VIRT, 
                RES=\RES,
                SHR=\SHR, 
                S=\S,
                CPU=\CPU,
                MEM=\MEM, 
                TIME=\TIME}
            {
                \PID & 
                \VIRT & 
                \RES &
                \SHR & 
                \S &
                \CPU &
                \MEM & 
                \TIME}
    \end{small}
\end{table}

\begin{lstlisting}[frame=single, style=shell]
    $ top -d 0.5 -p 1234 -b >> hasil-.txt    
\end{lstlisting}

\blindtext

\lstinputlisting[frame=single, style=python]{tables/performance-test/clean-file.py}