### 2021-03-26 include/kata-pengantar.tex
inc

\chapter*{KATA PENGANTAR}

\blindtext

\vspace{1cm}
\begin{flushright}
    Makassar, 9 April 2021\\
    \vspace{2.5cm}
    {SULAEMAN}\\
    NIM. {H13116002}
\end{flushright}